---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- This README is adapted from Josh London's repo (https://github.com/noaa-afsc/berchukseals-haulout/tree/main)-->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# put this under "badges: start" when zenodo DOI is available:
# [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4638221.svg)](https://doi.org/10.5281/zenodo.4638221)
```

<!-- badges: start -->

<!-- badges: end -->

## Ringed seal *(Pusa hispida)* haul-out behavior and emergence timing in the Bering, Chukchi, and Beaufort seas
Jessica M. Lindsay^1,2^,✉, Paul B. Conn^2^, Peter L. Boveng^2^, Justin A. Crawford^3^, Lori T. Quakenbush^3^, Andrew L. Von Duyke^4^, Kristin L. Laidre^1,5^

1. School of Aquatic and Fishery Sciences, University of Washington, Seattle, WA, USA 
2. Marine Mammal Laboratory, Alaska Fisheries Science Center, NOAA-NMFS, Seattle, WA, USA
3. Arctic Marine Mammal Program, Alaska Department of Fish and Game, Fairbanks, AK, USA
4. North Slope Borough Department of Wildlife Management, Utqiaġvik, AK, USA
5. Polar Science Center, Applied Physics Laboratory, University of Washington, Seattle, WA, USA

✉ Correspondence: Jessica M. Lindsay <jessica.lindsay@noaa.gov>

```{r, echo = FALSE}
# This repository serves as the research compendium in support of the above titled paper.
# As the manuscript works its way through peer review and the publications process,
# additional reproducibility features (e.g. devcontainer) and
# documentation of key revisions and releases
# 
# Major releases of the research compendium are published and archived with Zenodo [https://doi.org/10.5281/zenodo.4638221](https://doi.org/10.5281/zenodo.4638221)
```

## Draft Manuscript Under Active Development
This repository serves as the research compendium in support of the above titled paper, currently in review at Marine Ecology Progress Series.
Please note this manuscript is still in peer review and not final. Changes to code and results are still possible. As the manuscript works its way through peer review and the publications process, additional reproducibility features (e.g., Zenodo archive) will be added. A link to the published manuscript will be added here as well.

## Contents

This repository contains:

  - [:file\_folder: data](/data): cleaned haul-out and environmental data used for analysis
  
  - [:file\_folder: scripts](/scripts): scripts to run the analysis presented in the paper
  
  - [:file\_folder: figures](/figures): figures generated by the above scripts (Figs 3-6 and Figs S1-S3). Figures 1 (map) and 2 (conceptual diagram) in the manuscript were created outside of R. 

## Acknowledgements

We thank the Inuit communities, Brendan Kelly, and other field researchers who contributed to ringed seal tagging in the original studies. Tagging efforts were conducted under NMFS Research Permits Nos. 358–1787-01, 15324, and 20466 issued to ADF&G; 782-1694-00 issued to NOAA; and Scientific License Nos. SLE04/05-328 and SLE–05/06-322 from the Dept. of Fisheries and Oceans (DFO), Canada. Animal Care and Use Protocols were issued by the DFO (UFWI-ACC-2004-2005-001U) and the ADF&G Animal Care and Use Committee (Nos. 06–16, 2010-13R, 2014-03, 2016-23, 0027-2017-27, 0027-2018-29, and 0027-2019-41). Funding for tagging was provided by the Bureau of Ocean Energy Mgmt., the Office of Naval Research, the North Slope Borough (NSB)-Shell Baseline Studies Program, Shell Exploration and Production Co., National Fish and Wildlife Foundation with funding from Conoco Phillips, the Native Village of Kotzebue, NOAA, NMFS Alaska Region, and a U.S. Dept. of the Interior, Tribal Wildlife Grant for Federally Recognized Tribes. 

We also thank Stacie Koslovsky for assistance in accessing the ringed seal haul-out data and capture records, Brett McClintock for assistance with the momentuHMM R package, and Josh London for assistance with Argos location processing. We thank three anonymous reviewers whose feedback improved the manuscript. Support for Jessica Lindsay was provided by the National Science Foundation Graduate Research Fellowship Program under Grant No. DGE-1762114 and by the National Oceanic and Atmospheric Administration (NOAA). Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation or the National Marine Fisheries Service, NOAA. Any use of trade, firm, or product names does not imply endorsement by the U.S. Government. 
